---
title: "Lenz_B_NB_ReproductivePhysiology"
author: "Elizabeth Lenz"
date: "3/22/2019"
output: html_document
---

## Assessing impacts of bleaching susceptibility on sexual reproductive during recovery and sexual reproduction

# Packages required to run script

```{r setup, include=FALSE, message = FALSE}
rm(list=ls())
library(dplyr) #data manipulation
library(ggplot2) #Plotting package
library(MuMIn) #AIC package to select best fit model
library(effects) #model effects for anova (using lm): (plot(alleffects(model))))
library(car) #Anova(model): F tests
library(PerformanceAnalytics)
library(vegan)
library(psych)
library(gridExtra)
library(grid)
library(gtable)
library(Rmisc) #Basic statistical summary package
library(MASS)
library(aod) ## For Binary data
library(survival) #for Kaplan Meier Survivorship
library(devtools)
library(survminer) #for Kaplan Meier Survivorship
library(grid) #ggplotGrob 
library(pscl) #zero inflated
library(lme4) #linear mixed effects model
library(lmerTest)
library(emmeans) #posthoc analysis for lmer
library(cowplot) # grid.arrange axis alignment
library(glmmTMB) #glm for zero inflated data
library(rcompanion) #histogram plot residuals
library(DHARMa) # to inspect dispersion of glmer 
library(pbkrtest)
library(rstanarm)
library(tidyr)
library(betareg)

```

Setting working directory
```{r}
setwd("~/Desktop/Chapters/CH2_BNB_Repro/Rdata")
```

## Historical Temperatures
```{r}
Temp <- read.csv("NOAA.SWtemp2010_2016_MokuOLoe.csv", header = TRUE) #NOAA Buoy Mokuoloe Data from http://tidesandcurrents.noaa.gov/stationhome.html?id=1612480
Temp <- Temp[,c("DateTime","Temp")] # only select first two columns
Temp$DateTime <- as.character(Temp$DateTime) #to convert Date need to first format as character
Temp$DateTime <- as.POSIXct(Temp$DateTime, format = "%m/%e/%y %H:%M")#format date with month/day/year hour:minute
head(Temp)



Temp.Fig <- {plot(Temp~DateTime, data = Temp,ylab = "Temperature (°C)", col = "black", type = "l", lwd = 2, xlab = "Year", ylim = c(20,32)) #Plot NOAA Mokuloe Buoy data from 2014-2016
abline(h=30, lty = 2, lwd= 1, col = "darkred")
abline(h=31, lty = 2, lwd= 1, col = "darkred")} # Hortizontal line for 30C threshold

jpeg("temp.jpg")
Temp.Fig

while (!is.null(dev.list()))  dev.off()
```


## PART 1: IMPACTS OF BLEACHING ON REPRODUCTION

## SPAWNING DYNAMICS 
Collected from Bleached (n=8) and Nonbleached (n=7) colonies, watched during 5 night periods over 3 months of spawning in 2016 and 2017.

Proportion that spawned over night

```{r, echo = FALSE, warning = FALSE, message = FALSE}
spawn <- read.csv("Lenz_McapExSitu_Spawning2016_2017.csv")
spawn$SpawningMonth <- as.factor(spawn$SpawningMonth)
spawn$Year <- as.factor(spawn$Year)
spawn$Night <- as.factor(spawn$Night)
```

#GLM Model for Spawning Dynamics
```{r, echo = FALSE, warning = FALSE, message = FALSE}
spawn.glm <- glmer(Spawn ~ B.NB*Year + (1|Colony.ID) + (1|SpawningMonth), spawn, family = binomial) ###FIX!
Anova(spawn.glm, type = 3)

summary(spawn.glm)
shapiro.test(residuals(spawn.glm))
qqnorm(residuals(spawn.glm))
posthoc1 <- emmeans(spawn.glm, pairwise~Year, adjust = "tukey")
CLD(posthoc1, Letters=letters)
plot(allEffects(spawn.glm))

#spawn.table2 <- tab_model(spawn.glm, show.stat = TRUE, show.ci = 0.95, show.fstat = TRUE, show.se = TRUE, show.intercept = TRUE, string.se = "SE", string.stat = "Z-value",show.aic = TRUE, show.icc = TRUE, show.est = TRUE)

plot(allEffects(spawn.glm))
qqnorm(residuals(spawn.glm))
plot(residuals(spawn.glm))
simulationOutput <- simulateResiduals(fittedModel = spawn.glm)
plot(simulationOutput)
testUniformity(simulationOutput)
testDispersion(simulationOutput)
```

Create Figure for Spawning
```{r, echo = FALSE, warning = FALSE, message = FALSE}
spawnsuccess <- aggregate(Spawn~Colony.ID*Year*B.NB, data = spawn, sum, na.rm=TRUE)
write.csv(spawnsuccess, file = "spawncolony.csv")

spawn1 <- aggregate(Spawn~Year+B.NB+Date+SpawningMonth+Night, data = spawn, sum, na.rm=TRUE)

spawn1$totalcol <- ifelse(spawn1$B.NB == "Bleach", 8,
                          ifelse(spawn1$B.NB == "Nonbleach", 7, NA )) 
spawn1$percspawn <- spawn1$Spawn/spawn1$totalcol*100


spawn1$SpawningMonth <- as.numeric(spawn1$SpawningMonth)
spawn1$Date <- as.factor(spawn1$Date)
spawn1$MonthMoon <- paste(spawn1$SpawningMonth,spawn1$Night)
spawn1$propspawn <- spawn1$percspawn/100
spawn16 <- spawn1[ which(spawn1$Year==2016),] #subset 2016
spawn17 <- spawn1[ which(spawn1$Year==2017),] #subset 2017
  
spawn1$MonthMoon <- as.factor(spawn1$MonthMoon)
spawnprop.fig <- ggplot(data = spawn1, aes (x=MonthMoon, y = propspawn, group = B.NB, shape = B.NB)) +
  geom_line(aes(linetype = B.NB), size = 1) +
  geom_point(size=5, aes(color = B.NB), position = position_dodge(width = 0.5)) +
  scale_linetype_manual(values = c("dashed","solid")) +
  facet_grid(~Year) +
  scale_color_manual(values = c("Bleach" = "white", "Nonbleach" = "grey34")) +
  scale_shape_manual(values = c(19, 16)) +
  expand_limits(y=0) +
  ylab("Proportion Spawned") +
  xlab("Spawning Month") +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25), # panel.grid.minor=element_blank()
        plot.margin = margin(3,3,8,8, "pt"),
        panel.border = element_blank(), #  plot.background=element_blank()
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size=25), #Label Year       
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.key = element_blank(),
        legend.position = c(0.4,0.9),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15, angle = 90, hjust = 1),
        axis.title.x = element_text(size=20),
        axis.title.y = element_text(size=20))
ggsave("propspawnfig.pdf")
```

Proportion Spawned 
```{r, echo = FALSE, warning = FALSE, message = FALSE}
spawnsummary <- read.csv("spawnsummary.csv")
summary(spawnsummary)
spawnsummary$SpawningMonth <- as.factor(spawnsummary$SpawningMonth)

spawn2prop.fig <- ggplot(data = spawnsummary, aes (x=SpawningMonth, y = propspawn, colour = Parental.Performance, fill = Parental.Performance)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  facet_grid(~Year) +
  scale_fill_manual(values = c("Bleach" = "white", "Nonbleach" = "grey34")) +
  scale_shape_manual(values = c(19, 16)) +
  expand_limits(y=0) +
  ylab("Proportion Spawned") +
  xlab("Spawning Month") +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25), # panel.grid.minor=element_blank()
        plot.margin = margin(3,3,8,8, "pt"),
        panel.border = element_blank(), #  plot.background=element_blank()
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size=25), #Label Year       
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        legend.key = element_blank(),
        legend.position = c(0.5,0.9),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.x = element_text(size=20),
        axis.title.y = element_text(size=20))

spawn2prop.fig 
```

TOTAL REPRODUCTIVE OUTPUT
```{r, echo = FALSE, warning = FALSE, message = FALSE}

repro <- read.csv('Lenz_McapExSitu_ReproOutput2016-2017.csv') #create dataframe of reproductive output
summary(repro) #Summary of the units

hist(repro$microL.cm2, breaks = 30)
hist(log10(repro$microL.cm2+1)) 
repro$logOut <- log(repro$microL.cm2)  #Use log transformed data!
hist(repro$logOut)

repro$Year <- as.factor(repro$Year)
repro$SpawningMonth <- as.factor(repro$SpawningMonth)
repro$Colony <- as.factor(repro$Colony)
repro$Date <- as.character(repro$Date)
repro$Date <- as.Date(repro$Date, format = '%Y%m%d')
repro$DM <- format(repro$Date, format = "%m-%d")
repro$DM <- as.Date(repro$DM, format = "%m-%d")
```

Analyze fecundity over spawning seasons
```{r}

hist(repro$microL.cm2, breaks = 30)

rep1 <- aggregate(microL.cm2~Parental.Performance + Year + Colony, data = repro, sum, na.rm=TRUE)

repro.mod1 <- lmer(microL.cm2 ~ Parental.Performance*Year + (1|Colony), rep1)

Anova(repro.mod1, type =2)
summary(repro.mod1)
qqnorm(residuals(repro.mod1))
shapiro.test(residuals(repro.mod1))
TukeyHSD(aov(repro.mod1))

fec2 <- aggregate(microL.cm2~Parental.Performance + Colony, data = repro, sum, na.rm=TRUE)
hist(fec3$microL.cm2, breaks = 10)

repro.mod1 <- lm(microL.cm2 ~ Parental.Performance*Year, fec3)

Anova(repro.mod1, type = 2)
summary(repro.mod1)
qqnorm(residuals(repro.mod1))
shapiro.test(residuals(repro.mod1)) 

repro.mod2 <- lm(microL.cm2 ~ Parental.Performance, fec4)
Anova(repro.mod2, type =3)
TukeyHSD(repro.mod2)
summary(repro.mod2)
qqnorm(residuals(repro.mod2))
shapiro.test(residuals(repro.mod2))
#Visually inspect the residuals
shapiro.test(residuals(repro.mod1))
leveneTest(residuals(repro.mod1)~fec3$Parental.Performance)
boxplot(residuals(repro.mod1)~fec3$Parental.Performance)
qqnorm(residuals(repro.mod1))
plot(residuals(repro.mod1))
simulationOutput <- simulateResiduals(fittedModel = repro.mod1)
plot(simulationOutput)
testUniformity(simulationOutput)
testDispersion(simulationOutput)
```

```{r}

annfecsum <- summarySE(fec3, measurevar = "microL.cm2", groupvars = c("Year","Parental.Performance"))

annfec.fig <- ggplot(annfecsum, aes(x=Parental.Performance, y=microL.cm2)) + 
  geom_bar(position="dodge", stat = "identity", colour="black", aes(fill=Parental.Performance)) +
  geom_errorbar(aes(ymin=microL.cm2-se, ymax=microL.cm2+se), colour = "black", width = 0.2, position = position_dodge(0.9)) +
  facet_grid(~Year) + 
   scale_fill_manual(values = c("bisque", "chocolate4") ) +
  scale_y_continuous(expand = c(0,0)) +
  labs(y = expression(Fecundity~(~{mu}~L^{-2})), x = "Bleaching History") +
  coord_cartesian(ylim=c(0,30)) +
  #scale_fill_hue(name="", breaks = c("Bleached", "NonBleached"), labels = c("Bleached", "Nonbleached")) +
  #scale_fill_manual(values = c("white", "gray34")) +
  #ggtitle("A.") +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25),
        axis.title = element_text(size = 15),
        panel.grid.major=element_blank(),
        strip.text = element_text(size = 25),
        # panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        #  plot.background=element_blank(),
        legend.title=element_blank(),
        legend.text = element_blank(),
        legend.position = "none",
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_blank())
annfec.fig

annfec <- summarySE(fec3, measurevar = "microL.cm2", groupvars = "Parental.Performance")
(23.45-13.21)/23.45
(23.45-13.21)/((23.45+13.21)/2)

(23.82-18.45)/23.82*100
```


Fecundity: BleachingHistory / Month  / Parent
```{r}
mafecsum <- summarySE(fec2, measurevar = "microL.cm2", groupvars = c("Year","Parental.Performance", "SpawningMonth"))

ggplot(mafecsum, aes(x=SpawningMonth, y=microL.cm2, fill = Parental.Performance)) + 
  geom_bar(position="dodge", stat = "identity", colour="black") +
  facet_grid(~Year) +
  ggtitle("B.") +
  geom_errorbar(aes(ymin=microL.cm2-se, ymax=microL.cm2+se), colour = "black", width = 0.2, position = position_dodge(0.9)) +
   labs(y = expression(paste("Fecundity (", mu, "L cm"^{-2},")")), x = "Month - Day") +
  #coord_cartesian(ylim=c(0.375,0.7)) +
  xlab("Month") +
  scale_fill_hue(name="", breaks = c("Bleached", "NonBleached"), labels = c("Bleached", "Nonbleached")) +
  scale_fill_manual(values = c("gray92", "gray34")) 
```


Accumulative Curves for output:
```{r}
accu.fig <- ggplot(data = repro, aes(DM, AccumulativeOutput.mL.cm2, group=Colony, colour = Parental.Performance)) +
  geom_line() +
  geom_path() +
  geom_point() +
  labs(y = expression(Fecundity~(mL~cm^{-2})), x = "Month - Day") +
  ylim(0,100) + 
  facet_grid(~Year) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), # panel.grid.minor=element_blank()
        panel.border = element_blank(), #  plot.background=element_blank()
        legend.title=element_blank(),
        legend.position = c(0.2,0.8),
        legend.text = element_text(size=15),
        legend.key = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size=15),
        axis.title.x = element_text(size=20),
        strip.text = element_text(size = 20),
        strip.background = element_blank())
```

Fecundity for 2016 alone to inspect significance
```{r}

fec16 <- subset(fec3, Year == 2016)

fec16.lm <- aov(log(microL.cm2)~Parental.Performance, fec16)
Anova(fec16.lm)
summary(fec16.lm)
shapiro.test(residuals(fec16.lm))
qqnorm(residuals(fec16.lm))

fec16sum <- summarySE(fec16, measurevar = "microL.cm2", groupvars = c("Year","Parental.Performance"))

PerDiff16 <- (23.45-13.21)/(23.45)*100 #percent difference between parental bleaching history
PerDiff16

fec16.fig <- ggplot(fec16sum, aes(x=Parental.Performance, y=microL.cm2)) + 
  geom_bar(position="dodge", stat = "identity", colour="black", aes(fill=Parental.Performance)) +
  geom_errorbar(aes(ymin=microL.cm2-se, ymax=microL.cm2+se), colour = "black", width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = c("white", "gray34")) +
  labs(y = expression(Fecundity~(mL~cm^{-2})), x = "Bleaching History") +
  coord_cartesian(ylim=c(0,30)) +
  #scale_fill_hue(name="", breaks = c("Bleached", "NonBleached"), labels = c("Bleached", "Nonbleached")) +
  #scale_fill_manual(values = c("white", "gray34")) +
  #ggtitle("A.") +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25),
        axis.title = element_text(size = 15),
        panel.grid.major=element_blank(),
        # panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        #  plot.background=element_blank(),
        legend.title=element_blank(),
        legend.text = element_blank(),
        legend.position = "none",
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_text(size = 20))
fec16.fig

```

Fecundity for 2017 alone
```{r}

fec17 <- subset(fec3, Year == 2017)

fec17.lm <- aov((microL.cm2)~Parental.Performance, fec17)

Anova(fec17.lm)
shapiro.test(residuals(fec17.lm))
qqnorm(residuals(fec17.lm))

fec17sum <- summarySE(fec17, measurevar = "microL.cm2", groupvars = c("Year","Parental.Performance"))

PerDiff16 <- (23.45-13.21)/(23.45)*100 #percent difference between parental bleaching history
PerDiff16

fec17.fig <- ggplot(fec17sum, aes(x=Parental.Performance, y=microL.cm2)) + 
  geom_bar(position="dodge", stat = "identity", colour="black", aes(fill=Parental.Performance)) +
  geom_errorbar(aes(ymin=microL.cm2-se, ymax=microL.cm2+se), colour = "black", width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = c("white", "gray34")) +
  labs(y = expression(Fecundity~(mL~cm^{-2})), x = "Bleaching History") +
  coord_cartesian(ylim=c(0,30)) +
  scale_y_continuous(expand = c(0,0)) +
  #ggtitle("A.") +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25),
        axis.title = element_text(size = 15),
        panel.grid.major=element_blank(),
        # panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        #  plot.background=element_blank(),
        legend.title=element_blank(),
        legend.text = element_blank(),
        legend.position = "none",
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_text(size = 20))
fec17.fig

```

Egg Size between Parent Bleaching History x Year
```{r}
size <- read.csv("Lenz_McapExSitu_EggSize2016-2017.csv") #load in egg size csv
size <-na.omit(size) #remove NAs
size$SpawningMonth <- as.character(size$SpawningMonth)
size$Diameter.mm <- as.numeric(as.character(size$Diameter.mm)) #convert and maintain values
size$Diameter.micron <- size$Diameter.mm*1000
size$vol.mm3 <- as.character(size$vol.mm3)
size$vol.mm3 <- as.numeric(size$vol.mm3)
size$vol.microm3 <- (size$vol.mm3)*1000
hist(size$Diameter.micron, breaks = 20)
hist(size$vol.mm3, breaks = 20)

#### Assign colonies tag to correct Colony ID for individual ####
size$Colony <- ifelse(size$Colony.ID %in% c("146/326","146"), 146,
                          ifelse(size$Colony.ID %in% "117", 117, 
                          ifelse(size$Colony.ID %in% c("118/330", "118"), 118,
                          ifelse(size$Colony.ID %in% c("175/327", "175"), 175,
                          ifelse(size$Colony.ID %in% c("125/335", "125"), 125,             
                          ifelse(size$Colony.ID %in% c("116/332", "116"), 116, 
                          ifelse(size$Colony.ID %in% c("124/336", "124"), 124, 
                          ifelse(size$Colony.ID %in% c("149/329","149"), 149,              
                          ifelse(size$Colony.ID %in% c("173/331", "173"), 173, 
                          ifelse(size$Colony.ID %in% c("120/328", "120"), 120,                     
                          ifelse(size$Colony.ID %in% c("174/334", "174"), 174,
                          ifelse(size$Colony.ID %in% c("150/338", "150"), 150,
                          ifelse(size$Colony.ID %in% c("123/333", "123"), 123,
                          ifelse(size$Colony.ID %in% c("119/337", "119"), 119,
                          ifelse(size$Colony.ID %in% c("121/339", "121"), 121, NA))))))))))))))) #create new column with Original Colony Value
 
#### Assign Bleaching History to Colony ID ####

size$B.NB <- ifelse(size$Colony %in% c(146,118,116, 124, 120,174, 150), "Nonbleach",
                      ifelse(size$Colony %in% c(117,119, 175, 125, 149, 173, 123,121), "Bleach", NA))

size$Colony <- as.factor(size$Colony)
size$B.NB <- as.factor(size$B.NB)
size$SpawningMonth <- as.factor(size$SpawningMonth)
size$Date <- as.factor(size$Date)
size$Year <- as.factor(size$Year)
range(size$Diameter.micron)
range(size$vol.mm3)
```

```{r}


size <- size[which(size$Diameter.micron > 160),]

Diam <- summarySE(size, measurevar = "Diameter.micron", groupvars = c("Year", "SpawningMonth", "Colony", "B.NB", "Bundle.ID","Date"), na.rm=T)
range(Diam$Diameter.micron)

Diam1 <- summarySE(Diam, measurevar = "Diameter.micron", groupvars = c("Year", "SpawningMonth", "Colony", "B.NB","Date"), na.rm=T)

Diam2 <- summarySE(Diam1, measurevar = "Diameter.micron", groupvars = c("Year", "SpawningMonth", "Colony", "B.NB"), na.rm=T)

Diam3 <- summarySE(Diam2, measurevar = "Diameter.micron", groupvars = c("Year", "Colony", "B.NB"), na.rm=T)

Diam4 <- summarySE(Diam3, measurevar = "Diameter.micron", groupvars = c("Year", "B.NB"), na.rm=T)
```

Figure for Egg size by bleaching history and year
```{r}
EggDiam.fig <- ggplot(Diam4, aes(x=B.NB, y=Diameter.micron, fill = B.NB)) + 
  geom_bar(position="dodge", stat = "identity", colour="black") +
  scale_fill_manual(values = c("bisque", "chocolate4")) +
  facet_grid(~Year) +
  geom_errorbar(aes(ymin=Diameter.micron-se, ymax=Diameter.micron+se), colour = "black", width = 0.2) +
  coord_cartesian(ylim=c(400,480)) +
  ylab("Egg Diameter") +
  xlab("Bleaching History") +
  #scale_fill_hue(name="", breaks = c("Bleached", "NonBleached"), labels = c("Bleached", "Nonbleached")) +
   theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25),
        axis.title = element_text(size = 15),
        panel.grid.major=element_blank(),
        strip.text = element_blank(),
        # panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        #  plot.background=element_blank(),
        legend.title=element_blank(),
        legend.text = element_blank(),
        legend.position = "none",
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_text(size = 20)) 
EggDiam.fig  
```


Linear model Egg diameter ~ BleachingHistory * Year
```{r}
Diam.mod <- lmer(Diameter.micron ~ Year*B.NB + (1|Colony) + (1|SpawningMonth), Diam1)
Anova(Diam.mod)
summary(Diam.mod1)
qqnorm(residuals(Diam.mod))
shapiro.test(residuals(Diam.mod))

plot(residuals(Diam.mod))
simulationOutput <- simulateResiduals(fittedModel = Diam.mod)
plot(simulationOutput)
testUniformity(simulationOutput)
testDispersion(simulationOutput)
```

Egg volume ~ Year*BleachingHistory

```{r}
size1 <- size[which(size$vol.mm3 > 0.003),]
size1 <- size1[which(size$vol.mm3 < 0.16),]
hist(size1$vol.mm3,breaks = 50)

Vol <- summarySE(size, measurevar = "vol.mm3", groupvars = c("Year", "SpawningMonth", "Colony", "B.NB", "Bundle.ID","Date"), na.rm=T)
range(Vol$vol.mm3)
Vol <- na.omit(Vol)
vol.lm <- lm(Vol$vol.mm3~Vol$N)
Anova(vol.lm)
shapiro.test(residuals(vol.lm))
plot(Vol$vol.mm3~Vol$N)
abline(vol.lm)

Vol16 <- Vol[ which(Vol$Year==2016),] #subset 2016
Vol17 <- Vol[ which(Vol$Year==2017),] #subset 2016
B2016 <- Vol16[ which(Vol16$B.NB=="Bleach"),] #subset Bleach in 2016
B2017 <- Vol17[ which(Vol17$B.NB=="Bleach"),] #subset Bleach in  2017
NB2016 <- Vol16[ which(Vol16$B.NB=="Nonbleach"),] #subset Bleach in 2016
NB2017 <- Vol17[ which(Vol17$B.NB=="Nonbleach"),] #subset Bleach in  2017



lm <- lm(Vol$vol.mm3~Vol$N) #linear regression of egg size to lm
conf = seq(min(Diam$N),max(Diam$N),by = 0.05)

anova(lm)
summary(lm)

lm16 <- lm(Vol16$vol.mm3~Vol16$N)
anova(lm16)
summary(lm16)

lm17 <- lm(Vol17$vol.mm3~Vol17$N)
anova(lm17)
summary(lm17)

lmb16 <- lm(B2016$vol.mm3~B2016$N)
anova(lmb16)
summary(lmb16)

{plot(vol.mm3~N, Vol, ylab = "Egg Volume (mm3)", xlab = "Eggs per Bundle", col=ifelse(Vol$Year=="2016", "red", "blue"))

abline(lm, lwd = 2)
abline(lm16, lwd = 2, col = "red")
abline(lm17, lwd = 2, col = "blue")
legend(23, 0.57, legend=c("2016", "2017"),
       col=c("dark red", "dark blue"), lty = 1, lwd = 4, cex=2)}

abline(lmB16, col="red", lwd = 2)
abline(lmB17, col = "orange", lwd = 2)
abline(lmNB16, col = "blue", lwd = 2)
abline(lmNB17, col = "dark blue", lwd = 2)
abline(lm16, col = "dark red", lwd =4)
abline(lm17, col = "dark blue", lwd = 4)
legend(23, 0.57, legend=c("2016", "2017"),
       col=c("dark red", "dark blue"), lty = 1, lwd = 4, cex=2)

```
Vol1 <- summarySE(Vol, measurevar = "vol.mm3", groupvars = c("Year", "SpawningMonth", "Colony", "B.NB","Date"), na.rm=T)
range(Vol1$vol.mm3)


Vol2 <- summarySE(Vol1, measurevar = "vol.mm3", groupvars = c("Year", "SpawningMonth", "Colony", "B.NB"), na.rm=T)

Vol3 <- summarySE(Vol2, measurevar = "vol.mm3", groupvars = c("Year", "Colony", "B.NB"), na.rm=T)

Vol4 <- summarySE(Vol3, measurevar = "vol.mm3", groupvars = c("Year", "B.NB"), na.rm=T)
Vol4 <- Vol4[ c(1:4),]
Vol4
```

Figure for Egg Volume by Year * Bleaching History
```{r}
EggVol.fig <- ggplot(Vol4, aes(x=B.NB, y=vol.mm3, fill = B.NB)) + 
  geom_bar(position="dodge", stat = "identity", colour="black") +
  scale_fill_manual(values = c("bisque", "chocolate4")) +
  facet_grid(~Year) +
  geom_errorbar(aes(ymin=vol.mm3-se, ymax=vol.mm3+se), colour = "black", width = 0.2) +
  coord_cartesian(ylim=c(0.03,0.05)) +
  ylab("Egg Volume (mm3)") +
  xlab("Bleaching History") +
  #scale_fill_hue(name="", breaks = c("Bleached", "NonBleached"), labels = c("Bleached", "Nonbleached")) +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25),
        axis.title = element_text(size = 15),
        panel.grid.major=element_blank(),
        strip.text = element_text(size = 25),
        # panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        #  plot.background=element_blank(),
        legend.title=element_blank(),
        legend.text = element_blank(),
        legend.position = "none",
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_blank())
EggVol.fig  

```


```{r}
Vol.mod <- lmer(log10(vol.mm3) ~ Year*B.NB + (1|Colony) + (1|SpawningMonth),  Vol1)
Anova(Vol.mod, type = 2)
anova(Vol.mod, type =2 )
summary(Vol.mod)
qqnorm(residuals(Vol.mod))
shapiro.test(residuals(Vol.mod))
emmeans(Vol.mod, pairwise~Year*B.NB, adjust = "tukey")

plot(residuals(Vol.mod))
simulationOutput <- simulateResiduals(fittedModel = Vol.mod)
plot(simulationOutput)
testUniformity(simulationOutput)
testDispersion(simulationOutput)
```


Eggs per Bundle ~ BleachingHistory * Year
```{r}
buns <- read.csv("Lenz_McapExSitu_EggBundles2016-2017.csv")
buns <- na.omit(buns)
buns$Year <- as.factor(buns$Year)
range(buns$EggsPerBundle)
buns <- buns[which(buns$EggsPerBundle > 1),]
buns <- buns[which(buns$EggsPerBundle < 25),]
buns$sqEggsPerBundle <- sqrt(buns$EggsPerBundle)
hist(buns$EggsPerBundle)
range(buns$EggsPerBundle)

buns$B.NB <- ifelse(buns$Colony.ID %in% c(146,118,116, 124, 120,174, 150), "Nonbleach",
             ifelse(buns$Colony.ID %in% c(117,119, 175, 125, 149, 173, 123,121), "Bleach", NA)) #assign bleaching history to colony ID
buns$B.NB <- as.factor(buns$B.NB)

buns1 <- summarySE(buns, measurevar = "EggsPerBundle", groupvars = c("Year", "SpawningMonth", "Date", "Colony.ID", "B.NB"))
range(buns$EggsPerBundle)
range(buns1$EggsPerBundle)
buns2 <- summarySE(buns1, measurevar = "EggsPerBundle", groupvars = c("Year", "SpawningMonth","Colony.ID", "B.NB"))
range(buns2$EggsPerBundle)
buns3 <- summarySE(buns2, measurevar = "EggsPerBundle", groupvars = c("Year", "Colony.ID", "B.NB"))
buns4 <- summarySE(buns3, measurevar = "EggsPerBundle", groupvars = c("Year", "B.NB"))
buns5 <- summarySE(buns3, measurevar = "EggsPerBundle", groupvars = c("Year"))
(11.04-9.74)/(9.74)*100

```

Linear model
```{r}
hist(buns$EggsPerBundle)
buns.mod1 <- lmer((EggsPerBundle)~Year*B.NB + (1|Colony.ID) + (1|SpawningMonth), buns1)
Anova(buns.mod1, type = 2)
anova(buns.mod1, type = 2)
summary(buns.mod1)
emmeans(buns.mod1, pairwise~Year, adjust = "tukey")


qqnorm(residuals(buns.mod1))
plot(residuals(buns.mod1))
simulationOutput <- simulateResiduals(fittedModel = buns.mod1)
plot(simulationOutput)
testUniformity(simulationOutput)
testDispersion(simulationOutput)
shapiro.test(residuals(buns.mod1))
```

Buns Figure - Parent - Year

```{r}
Buns.fig <- ggplot(buns4, aes(x=B.NB, y=EggsPerBundle, fill = B.NB)) + 
  geom_bar(position="dodge", stat = "identity", colour="black") +
  scale_fill_manual(values = c("bisque", "chocolate4")) +
  facet_grid(~Year) +
  geom_errorbar(aes(ymin=EggsPerBundle-se, ymax=EggsPerBundle+se), colour = "black", width = 0.2, position = position_dodge(0.9)) +
  coord_cartesian(ylim=c(5,15)) +
  ylab("Eggs per Bundle") +
  xlab("Bleaching History") +
  #scale_fill_hue(name="", breaks = c("Bleached", "NonBleached"), labels = c("Bleached", "Nonbleached")) +
   theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25),
        axis.title = element_text(size = 15),
        panel.grid.major=element_blank(),
        strip.text = element_blank(),
        # panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        #  plot.background=element_blank(),
        legend.title=element_blank(),
        legend.text = element_blank(),
        legend.position = "none",
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_text(size = 20))
Buns.fig 
```


Percent Abnormal Eggs
```{r}
ab.buns <- summarySE(buns, measurevar = "Prop.Abnormal", groupvars = c("Year", "SpawningMonth", "Date","Colony.ID", "B.NB"), na.rm=T)
ab.buns$PerAb <- ab.buns$Prop.Abnormal*100
ab.buns1 <- summarySE(ab.buns, measurevar = "PerAb", groupvars = c("Year", "Colony.ID", "B.NB", "SpawningMonth"), na.rm=T)
ab.buns2 <- summarySE(ab.buns1, measurevar = "PerAb", groupvars = c("Year", "B.NB", "Colony.ID"), na.rm=T)
ab.buns3 <- summarySE(ab.buns2, measurevar = "PerAb", groupvars = c("Year", "B.NB"), na.rm=T)
ab.buns4 <- summarySE(ab.buns2, measurevar = "PerAb", groupvars = c("Year"), na.rm=T)
PerDiffAb <- (19.74-4.04)/19.74*100
PerDiffAb
```

Linear model to compare Percent Abnormal Eggs by bleaching history and year
```{r}
perab.lm <- lmer((PerAb)~Year*B.NB + (1|Colony.ID) + (1|SpawningMonth), ab.buns1)

Anova(perab.lm)
anova(perab.lm, type = 2)
summary(perab.lm)

boxplot(ab.buns$Prop.Abnormal~ab.buns$B.NB*ab.buns$Year)
plot(residuals(perab.lm))
qqnorm(residuals(perab.lm))
shapiro.test(residuals(perab.lm))

simulationOutput <- simulateResiduals(fittedModel = perab.lm)
plot(simulationOutput)
testUniformity(simulationOutput)
testDispersion(simulationOutput)
```


Percent Abnormal Eggs
```{r}
AbnormalEgg.fig <- ggplot(ab.buns2, aes(x=B.NB, y=PerAb)) + 
  geom_bar(position="dodge", stat = "identity", colour="black", aes (fill = B.NB)) +
  facet_grid(~Year) +
  geom_errorbar(aes(ymin=PerAb-se, ymax=PerAb+se), colour = "black", width = 0.2, position = position_dodge(0.9)) +
  labs(x = "Bleaching History", y = "% Egg Abnormality")  +
  coord_cartesian(ylim=c(0,30)) +
  scale_y_continuous(expand = c(0,0)) +
   scale_fill_manual(values = c("bisque", "chocolate4")) +
  #ggtitle("A.") +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25),
        axis.title = element_text(size = 15),
        panel.grid.major=element_blank(),
        strip.text = element_blank(),
        # panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        #  plot.background=element_blank(),
        legend.title=element_blank(),
        legend.text = element_blank(),
        legend.position = "none",
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_text(size = 20))

AbnormalEgg.fig
```

Combine graphics
```{r}
g <- ggplotGrob(spawnprop.fig)
g1 <- ggplotGrob(annfec.fig)
g2 <- ggplotGrob(EggVol.fig)
g3 <- ggplotGrob(Buns.fig)
g4 <- ggplotGrob(AbnormalEgg.fig)

Repro.Fig <- ggarrange(g1, g2, g3, g4, heights = c(2, 2),
          ncol = 2, nrow = 2, align = "v")


ggsave("ReproFitness.pdf", Repro.Fig, width = 8, height = 15, dpi = 1000)

while (!is.null(dev.list()))  dev.off()
```

## PART II: SELECTIVE BREEDING 

Attempted Crosses:
```{r}
BB <- 7/10*100
BNB <- 4/6*100
NBB <- 4/10*100
NBNB <- 7/8*100

c(BB,BNB,NBB, NBNB)
```


Fertilization Success
```{r}
fert <- read.csv("Mcap_FertilizationJuly2016.csv", header=TRUE, sep=",", na.strings=T)
read.csv("")
fert$perc.fert <- fert$Prop.Fert*100
fert$Dam <- as.factor(fert$Dam)
fert$Sire <- as.factor(fert$Sire)
head(fert)
summary(fert)

fert$SireBNB <- ifelse(fert$Sire %in% c(146,118,116, 124, 120,174, 150), "Nonbleach",
             ifelse(fert$Sire %in% c(117,119, 175, 125, 149, 173, 123,121), "Bleach", NA)) #assign bleaching history to colony ID

fert$DamBNB <- ifelse(fert$Dam %in% c(146,118,116, 124, 120,174, 150), "Nonbleach",
             ifelse(fert$Dam %in% c(117,119, 175, 125, 149, 173, 123,121), "Bleach", NA)) #assign bleaching history to colony ID

fert$SireBNB <- as.factor(fert$SireBNB)
fert$DamBNB <- as.factor(fert$DamBNB)
fert$Prop.2cell <- (fert$X2.cell/fert$Fertilized)
fert$Prop.4cell <- (fert$X4.cell/fert$Fertilized)
fert$Prop.8cell <- (fert$X8.cell/fert$Fertilized)
fert$Prop.16cell <- (fert$X16.cell/fert$Fertilized)


```


```{r}

plot(Prop.Fert~Cross,fert)

fert.sum <- summarySE(fert, measurevar = "perc.fert",groupvars = c("Cross","Conical","Dam","Sire", "DamBNB", "SireBNB"), na.rm=T)

aov()
fert.mod1 <- lmer(perc.fert~Cross + (1|Conical) + (1|Dam) + (1|Sire), fert)
fert.mod2 <- lmer(perc.fert~Cross + (1|Conical), fert)
anova(fert.mod1)
Anova(fert.mod1)
summary(fert.mod1)
emmeans(fert.mod1, pairwise~Cross, adjust = "Bonferroni") #TukeyHSD(fert.mod)
summary(glht(fert.mod1), linfct = mcp(Cross = "Tukey"))
summary(glht(fert.mod1, lsm(pairwise ~ Cross)))
TukeyHSD(aov(fert.mod1))

plotNormalHistogram(residuals(fert.mod1))

shapiro.test(residuals(fert.mod1))
simulationOutput <- simulateResiduals(fittedModel = fert.mod1)
plot(simulationOutput)
testUniformity(simulationOutput)
testDispersion(simulationOutput)

anova(fert.mod2)
Anova(fert.mod2)
summary(fert.mod2)
summary(glht(fert.mod2), linfct = mcp(Cross = "Tukey"))
summary(glht(fert.mod2, lsm(pairwise ~ Cross)))
emmeans(fert.mod2, pairwise~Cross, adjust = "Tukey") #TukeyHSD(fert.mod)

```
Fertilization Success
```{r}
fert.sum <- summarySE(fert, measurevar = "Prop.Fert",groupvars = c("Cross","Conical"), na.rm=T)

fert.sum2 <- summarySE(fert.sum, measurevar = "Prop.Fert", groupvars = "Cross", na.rm=T)
colnames(fert.sum2) <- c("Cross", "n", "mean", "sd", "sem", "ci")
fert.sum2$Cross<-c("BxB", "BxNB", "NBxB", "NBxNB")

CrossMean <- (59.33+39.75+38.58)/3
FertDiff <- ((84.57-CrossMean)/((84.57+CrossMean)/2))*100

Fert.fig <- ggplot(fert.sum2, aes(x=Cross, y = mean, sep = "")) + #plot objects
  geom_errorbar(aes(ymin=mean-sem, ymax=mean+sem), colour = "black", width = 0.2) + #add sem error bars
  geom_point(size = 8) + #include circular points
  scale_shape_manual(name = "", values = c(1,19)) + #B and NB symbol shapes
  ylab("Proportion Fertilized") +
  ylim(c(0,1)) +
  #ggtitle("Fertilization Success") +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold", size = 25),
        axis.title = element_text(size = 20),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border = element_blank(),
        plot.background=element_blank(),
        legend.position = "none",
        legend.key = element_blank(), 
        #legend.text = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.text = element_blank(),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 25),
        axis.text.x = element_text(size = 20),
        axis.title.x = element_text(size = 25))
Fert.fig

BB <- 8/10*100
BNB <- 4/6*100
NBB <- 4/7*100
NBNB <- 7/8*100

c(BB, BNB, NBB, NBNB)

```

#Developmental Stages

Analysis
```{r}

mod.2cell <- lmer(Prop.2cell~Cross + (1|Conical), fert)
fixef(mod.2cell)
Anova(mod.2cell)
anova(mod.2cell)
summary(mod.2cell)
qqnorm(residuals(mod.2cell))
plot(residuals(mod.2cell))
emmeans(mod.2cell, pairwise~Cross, adjust = "tukey")
shapiro.test(residuals(mod.2cell))

mod.4cell <- lmer(Prop.4cell~Cross + (1|Conical), fert)
Anova(mod.4cell)
anova(mod.4cell)
summary(mod.4cell)
emmeans(mod.4cell, pairwise~Cross, adjust = "tukey")
qqnorm(residuals(mod.4cell))
plot(residuals(mod.4cell))



mod.8cell <- lmer(Prop.8cell~Cross + (1|Conical), fert)
Anova(mod.8cell)
qqnorm(residuals(mod.8cell))
emmeans(mod.8cell, pairwise~Cross, adjust = "tukey")
plot(residuals(mod.4cell))
emmeans(mod.8cell, pairwise~Cross, adjust = "tukey")


mod.16cell <- lmer(Prop.16cell~Cross + (1|Conical),fert)
Anova(mod.16cell)
summary(mod.16cell)
qqnorm(residuals(mod.16cell))
plot(residuals(mod.16cell))
emmeans(mod.16cell, pairwise~Cross, adjust = "tukey")

```

```{r}
fert_long <- gather(fert, Devo, Prop, Prop.2cell, Prop.4cell, Prop.8cell, Prop.16cell, factor_key = TRUE)
head(fert_long)
fert_long <- na.omit(fert_long)

fert_long.ed <- subset(fert_long, select = c("Conical", "T1.Devo.Tub.No", "Cross", "Dam", "Sire", "SireBNB", "DamBNB", "Devo", "Prop"))
fert_long.ed$Prop <- as.numeric(fert_long.ed$Prop)
head(fert_long.ed)
dev.sum1 <- summarySE(fert_long.ed, measurevar = "Prop",groupvars = c("Cross","Conical", "SireBNB", "DamBNB", "Devo"), na.rm=T)

dev.sum2 <- summarySE(dev.sum1, measurevar = "Prop",groupvars = c("Cross", "SireBNB", "DamBNB", "Devo"), na.rm=T)
head(dev.sum2)

dev.sum2$newmean[dev.sum2$Devo == "Prop.16cell"] <- with(dev.sum2, Prop[Devo == "Prop.16cell"])
dev.sum2$newmean[dev.sum2$Devo == "Prop.8cell"] <- with(dev.sum2, Prop[Devo == "Prop.16cell"] + Prop[Devo == "Prop.8cell"])
dev.sum2$newmean[dev.sum2$Devo == "Prop.4cell"] <- with(dev.sum2, Prop[Devo == "Prop.16cell"] + Prop[Devo == "Prop.8cell"] + Prop[Devo == "Prop.4cell"])
dev.sum2$newmean[dev.sum2$Devo == "Prop.2cell"] <- with(dev.sum2, Prop[Devo == "Prop.16cell"] + Prop[Devo == "Prop.8cell"] + Prop[Devo == "Prop.4cell"] + Prop[Devo == "Prop.2cell"])
     
 
Dev.fig <- ggplot(data=dev.sum2, aes(x=Cross, y=Prop)) + 
  geom_bar(stat="identity", width = 0.4, aes(fill = Devo)) +
  geom_errorbar(aes(ymax=newmean+se,  ymin=newmean-se), width=0.15, position = "identity") +
  ylim(0,1.3) +
  xlab("Cross") + ylab("Proportion") + 
  scale_fill_manual(name="Developmental\nStage",
                         breaks=c("Prop.2cell", "Prop.4cell", "Prop.8cell", "Prop.16cell"),
                         labels=c("2-Cell", "4-Cell", "8-Cell", "16-Cell"),
                      values = c("lightgrey", "darkgrey", "dimgrey", "black")) +
  scale_x_discrete(limit = c("BB", "BNB", "NBB", "NBNB"),
                     labels = c("B×B","B×NB","NB×B", "NB×NB")) +
  theme(legend.position="right",
    axis.line.x = element_line(color = "black"),
    axis.line.y = element_line(color = "black"),
    axis.text.y = element_text(size = 20),
    axis.title.y = element_text(size = 25),
    axis.text.x = element_text(size = 20),
    axis.title.x = element_text(size = 25))
Dev.fig
```

```{r}
Embryo.Fig <- ggarrange(Fert.fig, Dev.fig, heights = c(3, 3),
          ncol = 1, nrow = 2, align = "h")

  
ggsave("Embryo.pdf", Embryo.Fig, width = 8, height = 15, dpi = 1000)
```

#Offspring: Larvae/Setters/Dead
```{r}
F1 <- read.csv("Lenz_Settlement3_CompData2016.csv")
F1$DayPostFert <- as.factor(F1$DayPostFert)

F1.larv <- subset(F1, select = c("Conical", "EggDonor", "SpermDonor", "Cross", "DayPostFert", "Prop.Larvae"))

larv.kt <- kruskal.test(Prop.Larvae~Cross, data=F1.larv)

larv.mod <- lmer(sqrt(Prop.Larvae) ~ Cross*DayPostFert + (1|EggDonor) + (1|SpermDonor), F1.larv)
Anova(larv.mod)
summary(larv.mod)
emmeans(larv.mod, pairwise~Cross*DayPostFert , adjust = "tukey")
shapiro.test(residuals(larv.mod))
qqnorm(residuals(larv.mod))


larv.mod2 <- glmer(sqrt(Prop.Larvae) ~ Cross*DayPostFert +  (1|Conical) + (1|EggDonor) + (1|SpermDonor), data= F1.larv, ziformula = ~1, family= poisson)


Anova(larv.mod2)
summary(larv.mod2)
shapiro.test(residuals(larv.mod2))
qqnorm(residuals(larv.mod2))

marginal = emtrends(larv.mod2, ~DayPostFert|Cross, var = "Cross")
pairs(marginal,adjust="tukey")
cld(marginal,
    alpha=0.05, 
    Letters=letters) 
plot(allEffects(larv.mod2,partial.residuals=TRUE))
plot(residuals(larv.mod2))



```

Larvae Plot
```{r}

F1.larv.sum <- summarySE(F1.larv, measurevar = "Prop.Larvae",groupvars = c("DayPostFert", "Cross"), na.rm=FALSE)


F1.larv.fig <- ggplot(data = F1.larv.sum, aes(x=DayPostFert, y = Prop.Larvae, group = Cross)) +
  geom_line(aes(color = Cross), size = 1, position = position_dodge(width = .2)) +
  geom_point(aes(color = Cross), size = 3, position = position_dodge(width = .2)) +
  geom_errorbar(aes(ymax=Prop.Larvae+se,  ymin=Prop.Larvae-se), width=0.15, position = position_dodge(width = .2)) +
  #scale_linetype_manual(values = c("dashed","solid")) +
  #scale_color_manual(values = c("Bleach" = "white", "Nonbleach" = "grey34")) +
  #scale_shape_manual(values = c(19, 16)) +
  ylab("Proportion Larvae") +
  xlab("Days Post-Fertilization") +
  scale_colour_manual(values = c("bisque", "darkgoldenrod3", "darkorange1", "chocolate4")) +
  #scale_colour_manual(values = col) +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25), # panel.grid.minor=element_blank()
        plot.margin = margin(3,3,8,8, "pt"),
        panel.border = element_blank(), #  plot.background=element_blank()
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size=25), #Label Year   
        legend.title = element_blank(),
        legend.text = element_blank(),
        legend.key = element_blank(),
        #legend.position = c(0.8,0.7),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20)) 
F1.larv.fig
```

```{r}
F1 <- read.csv("Lenz_Settlement3_CompData2016.csv")

F1.set <- subset(F1, select = c("Conical", "EggDonor", "SpermDonor", "Cross", "DayPostFert", "Prop.SettledAlive"))
F1.set$DayPostFert <- as.factor(F1.set$DayPostFert)

set.mod <- lmer(sqrt(Prop.SettledAlive) ~ Cross*DayPostFert + (1|EggDonor) + (1|SpermDonor), F1.set)
Anova(set.mod, type = 2)
anova(set.mod, type = 2)
summary(set.mod)
posthoc1 <- emmeans(set.mod, pairwise~Cross|DayPostFert, adjust = "tukey")
shapiro.test(residuals(set.mod))
qqnorm(residuals(set.mod))

set.mod2 <- glmer((Prop.SettledAlive) ~ Cross*DayPostFert + (1|EggDonor) + (1|SpermDonor) + (1|EggDonor:SpermDonor), family= poisson, F1.set)
Anova(set.mod2, type = 2)
anova(set.mod2, type = 2)
summary(set.mod2)
posthoc1 <- emmeans(set.mod, pairwise~Cross|DayPostFert, adjust = "tukey")
shapiro.test(residuals(set.mod))
qqnorm(residuals(set.mod))

F1.set.sum <- summarySE(F1.set, measurevar = "Prop.SettledAlive",groupvars = c("DayPostFert", "Cross"), na.rm=FALSE)


F1.set.fig <- ggplot(data = F1.set.sum, aes (x=DayPostFert, y = Prop.SettledAlive, group = Cross)) +
  geom_line(aes(color = Cross), size = 1, position = position_dodge(width = .2)) +
  geom_point(size=5, aes(color = Cross), position = position_dodge(width = .2)) +
  geom_errorbar(aes(ymax=Prop.SettledAlive+se,  ymin=Prop.SettledAlive-se), width=0.2, position = position_dodge(width = .2))  +
  scale_colour_manual(values = c("bisque", "darkgoldenrod3", "darkorange1", "chocolate4")) +
  ylab("Proportion Settled") +
  xlab("Days Post-Fertilization") +
  ylim(c(0,1)) +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25), # panel.grid.minor=element_blank()
        plot.margin = margin(3,3,8,8, "pt"),
        panel.border = element_blank(), #  plot.background=element_blank()
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size=25), #Label Year       
        legend.title = element_blank(),
        legend.text =element_blank(),
        legend.key = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20))
F1.set.fig
```

```{r}
F1.mort <- subset(F1, select = c("Conical", "EggDonor", "SpermDonor", "Cross", "DayPostFert", "Dead"))

mort.lm <- lm(Dead~Cross*DayPostFert, F1.mort)
summary(aov(mort.lm))

mort.mod <- lmer((Dead) ~ Cross + (1|DayPostFert), F1.mort)
Anova(mort.mod)
anova(mort.mod)
summary(mort.mod)
shapiro.test(residuals(mort.mod))
qqnorm(residuals(mort.mod))

F1.mort.sum <- summarySE(F1.mort, measurevar = "Dead", groupvars = c("DayPostFert", "Cross"), na.rm=FALSE)

F1.mort.fig <- ggplot(data = F1.mort.sum, aes (x=DayPostFert, y = Dead, group = Cross)) +
  geom_line(aes(color = Cross), size = 1,  position = position_dodge(width = 2)) +
  geom_point(size=5, aes(color = Cross), position = position_dodge(width = 2)) +
  geom_errorbar(aes(ymax=Dead+se,  ymin=Dead-se), width=0.2, position = position_dodge(width = 2)) +
  scale_colour_manual(values = c("bisque", "darkgoldenrod3", "darkorange1", "chocolate4")) +
  ylab("Proportion Mortality") +
  xlab("Days Post-Fertilization") +
  ylim(c(0,1)) +
  theme_bw() +
  theme(plot.title = element_text(lineheight = 0.8, face = "bold",size = 25), # panel.grid.minor=element_blank()
        plot.margin = margin(3,3,8,8, "pt"),
        panel.border = element_blank(), #  plot.background=element_blank()
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text = element_text(size=25), #Label Year 
        legend.title = element_blank(),
        legend.text =element_blank(),
        legend.key = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15, angle = 90, hjust = 1),
        axis.title.x = element_text(size=20),
        axis.title.y = element_text(size=20))
F1.mort.fig
```


```{r}
pdf('F1.phy2016.pdf', width=8, height=10)
grid.arrange((arrangeGrob(F1.larv.fig, F1.set.fig, cross.km, nrow=3)))
dev.off()

p1 <- ggplotGrob(F1.larv.fig)
p2 <- ggplotGrob(F1.set.fig)
p3 <- ggplotGrob(cross.km)

F1.Fig <- ggarrange(Fert.fig, F1.larv.fig, F1.set.fig,cross.km, heights = c(3, 3),
          ncol = 2, nrow = 3, align = "v")

  
ggsave("F1.pdf", F1.Fig, width = 8, height = 15, dpi = 1000)
dev.off()
while (!is.null(dev.list()))  
```




Kaplan Meier Survivorship Estimate
```{r}

km <-read.csv("Lenz_KM_SurvivalTest2016.csv", header=TRUE, sep=",", na.strings="NA")

km$status <- ifelse(km$censor %in% 0, 1,
                   ifelse(km$censor %in% 1, 2 ,NA)) 

km$crossN <- ifelse(km$Cross %in% "BB", 1,
                   ifelse(km$Cross %in% "BNB", 2,
                          ifelse(km$Cross %in% "NBB", 3,
                                 ifelse(km$Cross %in% "NBNB", 4,NA))))

cox.crossN <- coxph(Surv(Timepoint,status) ~ crossN, data = km) #create model
summary(cox.crossN) # statistical summary 

cox.cross <- coxph(Surv(Timepoint,status) ~ Cross, data = km) #create model
summary(cox.cross) # statistical summary 

res <- pairwise_survdiff(Surv(Timepoint, status) ~ Cross*Conical.ID,
     data = km)
symnum(res$p.value, cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
   symbols = c("****", "***", "**", "*", "+", " "),
   abbr.colnames = FALSE, na = "")

km$crossD <- ifelse(km$Cross %in% c("BB","BNB"), "B",
                   ifelse(km$Cross %in% c("NBB", "NBNB"), "NB", NA))

cox.crossD <- coxph(Surv(Timepoint,status) ~ crossD, data = km) #create model
summary(cox.crossD) # statistical summary  
                     
km$crossS <- ifelse(km$Cross %in% c("BB","NBB"), "B",
                   ifelse(km$Cross %in% c("BNB", "NBB"), "NB", NA))

cox.crossS <- coxph(Surv(Timepoint,status) ~ crossS, data = km) #create model
summary(cox.crossS) # statistical summary 

cox.zph(cox.crossN) 
cox.zph(cox.crossD)
cox.zph(cox.crossS) #assumptions of statistical analyses

```

```{r}
cox.egg <- coxph(Surv(Timepoint,censor) ~ Egg, km)
summary(cox.egg)

cox.sperm <- coxph(Surv(Timepoint,censor) ~ Sperm, km)
summary(cox.sperm)

cross.km <- ggsurvplot(cox.cross,              # survfit object with calculated statistics.
                       risk.table = FALSE,       # show risk table.
                       pval = TRUE,             # show p-value of log-rank test.
                       pval.size=10,
                       conf.int = TRUE,         # show confidence intervals for point estimaes of survival curves.
                       main="Cox Model",
                       linetype = "strata",
                       size = 1,
                       font.x = c(18, "bold", "black"),
                       font.y = c(18, "bold", "black"), 
                       palette = c("burlywood2", "chocolate3"),
                       legend.title = "Cross",
                       legend.labs = c("BxB", "BxNB", "NBxB", "NBxNB"),
                       xlab = "",
                       ylab = "",
                       font.main = c(20, "bold", "black"),
                       font.legend = c(16, "bold", "black"))





cross.fit <- survfit(Surv(Timepoint, censor) ~ + Cross, type="kaplan-meier", conf.type="log", data = km)
summary(cross.fit)

BB.fit <- (0.93-0.07)/0.93*100
BNB.fit <- (0.955-0.117)/0.955*100
NBB.fit <- (0.994-0.143)/0.994*100
NBNB.fit <- (0.947-0.215)/0.947*100

(0.215-0.07)/0.215*100
(0.143-0.07)/0.142*100

((67.44 + 51.40)/2)



cross.km <- ggsurvplot(cross.fit,              # survfit object with calculated statistics.
                    risk.table = FALSE,       # show risk table.
                    pval = TRUE,             # show p-value of log-rank test.
                    #pval.size=10,
                    #conf.int = TRUE,         # show confidence intervals for point estimaes of survival curves
                    main="Kaplan-Meier Analysis",
                    #linetype = "strata",
                    size = 2,
                    font.x = c(18, "bold", "black"),
                    font.y = c(18, "bold", "black"), 
                    legend.title = "Cross",
                    legend.labs = c("B x B", "B x NB", "NB x B", "NB x NB"),
                    palette = c("bisque", "darkgoldenrod3","darkorange1", "chocolate4"),
                    xlab = "",
                    ylab = "",
                    font.main = c(20, "bold", "black"),
                    font.legend = c(16, "bold", "black"))

cross.km
```

```{r}
egg.fit <- survfit(Surv(Timepoint, censor) ~ Egg, type="kaplan-meier", conf.type="log", data = km)
summary(egg.fit)
egg.km <- ggsurvplot(egg.fit,                     # survfit object with calculated statistics.
  risk.table = FALSE,       # show risk table.
  size = 2,
  pval = TRUE,             # show p-value of log-rank test.
  pval.size=10,
  #conf.int = TRUE ,         # show confidence intervals for point estimaes of survival curves.
  font.x = c(18, "bold", "black"),
  font.y = c(18, "bold", "black"), 
  font.main = c(20, "bold", "black"),
  font.legend = c(16, "bold", "black"),
  palette = c("bisque", "chocolate4"),
  xlab = "",
  legend.title = "Dam",
  legend.labs = c("Bleached", "Nonbleached")) #present narrower X axis, but not affect survival estimates.

egg.km
```

```{r}
sperm.fit <- survfit(Surv(Timepoint, censor) ~ Sperm, type="kaplan-meier", conf.type="log", data = km)
summary(sperm.fit)

sperm.km <- ggsurvplot(sperm.fit,                     # survfit object with calculated statistics.
                     size = 2,
                     #risk.table = TRUE,       # show risk table.
                     #risk.table.y.text.col = TRUE,
                     pval = TRUE,             # show p-value of log-rank test.
                     pval.size=10,
                     #break.time.by = c(6, 7, 28, 53, 59),
                     #tables.height = 0.2,
                     #tables.theme = theme_cleantable(),
                     #conf.int = TRUE ,         # show confidence intervals for point estimaes of survival curves.
                     #conf.inf.style = c("ribbon","step"),
                     xlab = "Days Post-Fertilization",
                     ylab = "",
                     font.x = c(18, "bold", "black"),
                     font.y = c(18, "bold", "black"), 
                     font.main = c(20, "bold", "black"),
                     palette = c("bisque", "chocolate4"),
                     font.legend = c(16, "bold", "black"),
                     legend.title = "Sire",
                     legend.labs = c("Bleached", "Nonbleached")) # present narrower X axis, but not affect survival estimates.

sperm.km

 tables <- ggsurvtable(sperm.fit, data = Surv, color = "strata",
      y.text = FALSE)
    # Risk table
    tables$risk.table
    # Number of cumulative events
    tables$cumevents
    # Number of cumulative censoring
    tables$cumcensor
```

```{r}

splots <- list(cross.km, egg.km, sperm.km)

res <- arrange_ggsurvplots(splots, print = FALSE, ncol=1, nrow =3, risktable.height = 0.4)
ggsave("F1_KM2.pdf", res, width = 8, height = 15, dpi = 1000)
dev.off()

```

